### 2018-12-23 tim-lucas-wsc.tex
Sort refs out I think.

@@

\begin{center}
{Tim C. D. Lucas\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
{1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\

@@

High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \cite{sturrock2016mapping, sturrock2014fine}. 
Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}. 
However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
Given these restrictions, modelling any of the non linearities known to be important in malaria estimation is difficult. 
A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.

One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data. 
This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{sill2009feature, bhatt2017improved}. 
Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
As case studies we use data from Colombia and Madagascar as these countries have  good routine surveillance data.

@@

The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
For Colombia we used all points from South America while for Madagascar we used only Madagascan data.

The polygon level surveillance data were collated from various government reports. To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \cite{cibulskis2011worldwide} were used. Where species specific reports were given, these were used. For reports where no species specific, national estimates of the ratio between \emph{P. falciparum} and \emph{P. vivax} cases were used to calculate \emph{P. falciparum} only cases. To minimise temporal effects we selected, for each country, one year of surveillance data. 
We used surveillance data from 2015 for Colombia and data from 2013 for Indonesia.

Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \cite{gpw4} and WorldPop \cite{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data. 
Final population rasters were created by linear interpolation of the surrounding five-yearly rasters. 
Finally, for each year, national population estimates from the UN were raked over the interpolated population surfaces.

@@

For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{wright2015ranger}, projection pursuit regression \cite{friedman1981projection}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
%The parameters tried can be seen in Table S1.
Predictions from these models were then made across Colombia and Madagascar.
These predictions were finally inverse logit transformed so that they are on the correct scale.

The top level model is a disaggregation regression model \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.

@@

where $\mathrm{p2i}$ is from a model that was published previously \cite{cameron2015defining} which defines a function
$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$

@@

Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.

We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}. We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}

@@

\begin{figure}
\centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
        \caption{Random cross-validation}
        \label{fig:cv1}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
        \caption{Spatial cross-validation}
        \label{fig:cv1}
    \end{subfigure}
\caption{
  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
  a) Six-fold random cross-validation.
  b) Three-fold spatial cross-validation with folds indicated by colour.
}
\label{f:scatter}
\end{figure}

@@

{\bf References}\\

\bibliography{Malaria} 
\bibliographystyle{plain}

>>>>

\usepackage{natbib}

%\setcitestyle{authoryear}

@@

\begin{center}
{Tim C. D. Lucas*\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
{1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\

@@

High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}. 
Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}. 
However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
Given these restrictions, modelling any of the non linearities known to be important in malaria estimation is difficult. 
A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.

One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data. 
This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \citep{sill2009feature, bhatt2017improved}. 
Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
As case studies we use data from Colombia and Madagascar as these countries have  good routine surveillance data.

@@

The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \citep{bhatt2015effect}. 
As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \citep{smith2007standardizing}. 
For Colombia we used all points from South America while for Madagascar we used only Madagascan data.

The polygon level surveillance data were collated from various government reports. To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \citep{cibulskis2011worldwide} were used. Where species specific reports were given, these were used. For reports where no species specific, national estimates of the ratio between \emph{P. falciparum} and \emph{P. vivax} cases were used to calculate \emph{P. falciparum} only cases. To minimise temporal effects we selected, for each country, one year of surveillance data. 
We used surveillance data from 2015 for Colombia and data from 2013 for Indonesia.

Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \citep{gpw4} and WorldPop \citep{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data. 
Final population rasters were created by linear interpolation of the surrounding five-yearly rasters. 
Finally, for each year, national population estimates from the UN were raked over the interpolated population surfaces.

@@

For each region we fitted five models via caret \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
%The parameters tried can be seen in Table S1.
Predictions from these models were then made across Colombia and Madagascar.
These predictions were finally inverse logit transformed so that they are on the correct scale.

The top level model is a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.

@@

where $\mathrm{p2i}$ is from a model that was published previously \citep{cameron2015defining} which defines a function
$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$

@@

Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \citep{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.

We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}. We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}

@@

\begin{figure}
\centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
        \caption{Random cross-validation}
        \label{fig:cv1}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
        \caption{Spatial cross-validation}
        \label{fig:cv2}
    \end{subfigure}
\caption{
  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
  a) Six-fold random cross-validation.
  b) Three-fold spatial cross-validation with folds indicated by colour.
}
\label{f:scatter}
\end{figure}

@@

\bibliography{Malaria} 
\bibliographystyle{apalike}
